---
title: "Reproducible Research - PA1_template"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Q1 Loading and preprocessing the data

```{r}
Sys.setlocale("LC_TIME", "English")

if(!file.exists("activity.csv")) {
  temp <- tempfile()
  download.file("http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",temp)
  unzip(temp)
  unlink(temp)
}
activity <- read.csv("activity.csv")
summary(activity)
```
## Q2 What is the average daily activity pattern

```{r,message=FALSE, warning=FALSE}
library(lubridate)
```
```{r}
activity$date <- ymd(activity$date)
str(activity)
```

```{r ,message=FALSE, warning=FALSE}
require(dplyr)
```
```{r}
totalStepsPerDay <- activity %>% group_by(date) %>%summarise(totalSteps=sum(steps),na=sum(is.na(steps))) %>% print
hist(totalStepsPerDay$totalSteps, main =paste("Total Steps Each Day"), xlab="Number of Steps",col="green")
```

```{r}
meanSteps <- mean(totalStepsPerDay$totalSteps,na.rm=TRUE)
medianSteps <- median(totalStepsPerDay$totalSteps,na.rm=TRUE)
```
```{r}
meanSteps
medianSteps
```
#Q3 What is the average daily activity pattern

```{r ,message=FALSE, warning=FALSE}
library(ggplot2)
```
```{r}
stepsByInterval <- activity%>%group_by(interval)%>%summarise(average_steps=mean(steps, na.rm=TRUE))
head(stepsByInterval)

ggplot(stepsByInterval, aes(x =interval , y=average_steps)) +
  geom_line(color="orange", size=1) +
  labs(title = "Average Daily Activity", x = "Interval", y = "Average Steps Across All Day")
```

```{r}
maxInterval <- stepsByInterval[which.max(stepsByInterval$average_steps),]
maxInterval
```
## Q4 Imputing missing values

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)##

```{r}
sum(is.na(activity$steps))
```
2. Devise a strategy for filling in all of the missing values in the dataset.

Strategy: filling in missing data with the mean number of steps across all days with available data for that particular interval.

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
StepsPerInterval <- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)

activitySplit <- split(activity, activity$interval)

for(i in 1:length(activitySplit)){
  activitySplit[[i]]$steps[is.na(activitySplit[[i]]$steps)] <- StepsPerInterval[i]
}
activityNoNas <- do.call("rbind", activitySplit)
activityNoNas <- activityNoNas[order(activityNoNas$date),]

```

4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.

```{r}
StepsPerDayNoNas<- tapply(activityNoNas$steps, activityNoNas$date, sum)

hist(StepsPerDayNoNas, xlab = "Number of Steps", main = "Steps per Day (without NA)", col="orange")
```
```{r}
MeanPerDayNoNas <- mean(StepsPerDayNoNas, na.rm = TRUE)
MeanPerDayNoNas
MedianPerDayNoNas <- median(StepsPerDayNoNas, na.rm = TRUE)
MedianPerDayNoNas
```


## Q5 Are there differences in activity patterns between weekdays and weekends?


1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}
activityNoNas$day <- ifelse(weekdays(as.Date(activityNoNas$date),abbreviate = FALSE) == "Saturday" | weekdays(as.Date(activityNoNas$date),abbreviate = FALSE) == "Sunday", "weekend", "weekday")
```

2.Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)

```{r}
StepsPerIntervalWeekend <- tapply(activityNoNas[activityNoNas$day == "weekend" ,]$steps, activityNoNas[activityNoNas$day == "weekend" ,]$interval, mean, na.rm = TRUE)

StepsPerIntervalWeekday <- tapply(activityNoNas[activityNoNas$day == "weekday" ,]$steps, activityNoNas[activityNoNas$day == "weekday" ,]$interval, mean, na.rm = TRUE)
```
Plots of weekdays and weekend activity
```{r}
par(mfrow=c(1,2))
#Weekend activity - plot
plot(as.numeric(names(StepsPerIntervalWeekend)), 
     StepsPerIntervalWeekend, 
     xlab = "Interval", 
     ylab = "Steps", 
     main = "Activity Pattern (Weekends)", 
     type = "l")
#Weekday activity - plot
plot(as.numeric(names(StepsPerIntervalWeekday)), 
     StepsPerIntervalWeekday, 
     xlab = "Interval", 
     ylab = "Steps", 
     main = "Activity Pattern (Weekdays)", 
     type = "l")
```





